---
title: "R Notebook"
output: html_notebook
---

```{r}
library(readxl)
library(tidyverse)
```

```{r}
df <- read_xlsx('data/PFIII_ex1_2020_data.xlsx')
```


```{r}
LL_1 <- -4271.1
LL_2 <- -4257.4
k <- 5
AIC_1 <- 2 * k - 2 * LL_1
AIC_2 <- 2 * k - 2 * LL_2

rho_2 <- 1 - (LL_2/LL_1)
```

Based on the estimated parameters, calculate the individual probabilities for each alternative in each observation. Compute the market shares for each mode in the base situation and compare to the observed market shares.

```{r}
alpha <- -0.4489316
beta <- -0.0267934
v1 <- alpha*log(df$TC1) + beta*df$TT1
v2 <- 0.1097849 + alpha*log(df$TC2) + beta*df$TT2
v4 <- 1.3413546 + alpha*log(df$TC4) + beta*df$TT4
v6 <- -0.7295991 + alpha*log(df$TC6) + beta*df$TT6
```

```{r}
df$prob_1 <- exp(v1) / (exp(v1) + exp(v2) + exp(v4) + exp(v6))
df$prob_2 <- exp(v2) / (exp(v1) + exp(v2) + exp(v4) + exp(v6))
df$prob_4 <- exp(v4) / (exp(v1) + exp(v2) + exp(v4) + exp(v6))
df$prob_6 <- exp(v6) / (exp(v1) + exp(v2) + exp(v4) + exp(v6))
```

```{r}
select(df, c("prob_1","prob_2","prob_4","prob_6"))
```

```{r}
c(mean(df$prob_1),mean(df$prob_2),mean(df$prob_4),mean(df$prob_6))
```


```{r}
estimations <- data.frame("trips"=c(mean(df$prob_1)*4197, mean(df$prob_2)*4197, mean(df$prob_4)*4197, mean(df$prob_6)*4197), "mode"=c("walk","bike","car","public transport"))
```

```{r}
sum(estimations$trips)
```

#Create better histograms

```{r}
table(df$choice)
```



```{r}
barplot(height=rle(sort(df$choice))$lengths, names=estimations$mode, col="cyan3")
```

```{r}
barplot(height=estimations$trips, names=estimations$mode, col="lightslateblue")
```
```{r}
#create dataframe
barframe <- data.frame("trips"=c(estimations$trips, rle(sort(df$choice))$lengths), "mode"=rep(estimations$mode,2), "method"=rep(c("model","empirical"), each=4))

# Modify data for Base R barplot
barframe <- reshape(barframe,                        
                     idvar = "method",
                     timevar = "mode",
                     direction = "wide")
row.names(barframe) <- barframe$method
barframe <- barframe[ , 2:ncol(barframe)]
colnames(barframe) <- estimations$mode
barframe <- as.matrix(barframe)

# Grouped barplot using Base R
barplot(height = barframe, beside = TRUE, col=c("lightslateblue", "cyan3"))
legend(1,2000, legend=c("model","empirical"), fill=c("lightslateblue", "cyan3"))                 
```
```{r}
df$VOT_walk <- beta / (alpha/df$TC1)
df$VOT_bike <- beta / (alpha/df$TC2)
df$VOT_car <- beta / (alpha/df$TC4)
df$VOT_pub <- beta / (alpha/df$TC6)
mean(df$VOT_walk)
mean(df$VOT_bike)
mean(df$VOT_car)
mean(df$VOT_pub)
```
```{r}
sum(df$VOT_walk * df$prob_1) / sum(df$prob_1)
sum(df$VOT_bike * df$prob_2) / sum(df$prob_2)
sum(df$VOT_car * df$prob_4) / sum(df$prob_4)
sum(df$VOT_pub * df$prob_6) / sum(df$prob_6)
```






Now, you would like to investigate whether there is a difference in preferences for using public
transport between male and female respondents. Make a cross table presenting how the market
shares are split across gender.
In the output presented below, the interaction variable Ptfem has been included, which is only
one in the public transport alternative and only if the user is female:

```{r}
df_fem=df
df_fem$Gender=df_fem$Gender-1
```


```{r}
alpha <- -0.4747656
beta <- -0.0263811
c=0.9936726
v1 <- alpha*log(df_fem$TC1) + beta*df_fem$TT1
v2 <- 0.1323373 + alpha*log(df_fem$TC2) + beta*df_fem$TT2
v4 <- 1.4104197 + alpha*log(df_fem$TC4) + beta*df_fem$TT4
v6 <- -1.2657003 + alpha*log(df_fem$TC6) + beta*df_fem$TT6+c*df_fem$Gender
```


```{r}
df_fem$prob_1 <- exp(v1) / (exp(v1) + exp(v2) + exp(v4) + exp(v6))
df_fem$prob_2 <- exp(v2) / (exp(v1) + exp(v2) + exp(v4) + exp(v6))
df_fem$prob_4 <- exp(v4) / (exp(v1) + exp(v2) + exp(v4) + exp(v6))
df_fem$prob_6 <- exp(v6) / (exp(v1) + exp(v2) + exp(v4) + exp(v6))
```


```{r}
c(
  mean(df$prob_1),
  mean(df$prob_2),
  mean(df$prob_4),
  mean(df$prob_6))
```
```{r}
c(
  mean(df_fem$prob_1),
  mean(df_fem$prob_2),
  mean(df_fem$prob_4),
  mean(df_fem$prob_6))
```



```{r}
df_fem_split=filter(df_fem, Gender==1)
df_male_split=filter(df_fem, Gender==0)
```

```{r}
males_walk <- mean(df_male_split$prob_1)*nrow(df_male_split)
females_walk <- mean(df_fem_split$prob_1)*nrow(df_fem_split)
females_walk / (males_walk + females_walk)

males_bike <- mean(df_male_split$prob_2)*nrow(df_male_split)
females_bike <- mean(df_fem_split$prob_2)*nrow(df_fem_split)
females_bike / (males_bike + females_bike)

males_car <- mean(df_male_split$prob_4)*nrow(df_male_split)
females_car <- mean(df_fem_split$prob_4)*nrow(df_fem_split)
females_car / (males_car + females_car)

males_pt <- mean(df_male_split$prob_6)*nrow(df_male_split)
females_pt <- mean(df_fem_split$prob_6)*nrow(df_fem_split)
females_pt / (males_pt + females_pt)
```

```{r}
df_fem_split=filter(df, Gender==2)
df_male_split=filter(df, Gender==1)
```

```{r}
males_walk <- mean(df_male_split$prob_1)*nrow(df_male_split)
females_walk <- mean(df_fem_split$prob_1)*nrow(df_fem_split)
females_walk / (males_walk + females_walk)

males_bike <- mean(df_male_split$prob_2)*nrow(df_male_split)
females_bike <- mean(df_fem_split$prob_2)*nrow(df_fem_split)
females_bike / (males_bike + females_bike)

males_car <- mean(df_male_split$prob_4)*nrow(df_male_split)
females_car <- mean(df_fem_split$prob_4)*nrow(df_fem_split)
females_car / (males_car + females_car)

males_pt <- mean(df_male_split$prob_6)*nrow(df_male_split)
females_pt <- mean(df_fem_split$prob_6)*nrow(df_fem_split)
females_pt / (males_pt + females_pt)
```



```{r}
males_walk <- nrow(filter(df_male_split, Mode == 1))
females_walk <- nrow(filter(df_fem_split, Mode == 1))
females_walk / (males_walk + females_walk)

males_bike <- nrow(filter(df_male_split, Mode == 2))
females_bike <- nrow(filter(df_fem_split, Mode == 2))
females_bike / (males_bike + females_bike)

males_car <- nrow(filter(df_male_split, Mode == 4))
females_car <- nrow(filter(df_fem_split, Mode == 4))
females_car / (males_car + females_car)

males_pt <- nrow(filter(df_male_split, Mode == 6))
females_pt <- nrow(filter(df_fem_split, Mode == 6))
females_pt / (males_pt + females_pt)
```





```{r}
df$TC4 <- df$TC4*1.50
alpha <- -0.4489316
beta <- -0.0267934
v1 <- alpha*log(df$TC1) + beta*df$TT1
v2 <- 0.1097849 + alpha*log(df$TC2) + beta*df$TT2
v4 <- 1.3413546 + alpha*log(df$TC4) + beta*df$TT4
v6 <- -0.7295991 + alpha*log(df$TC6) + beta*df$TT6
```

```{r}
df$prob_1 <- exp(v1) / (exp(v1) + exp(v2) + exp(v4) + exp(v6))
df$prob_2 <- exp(v2) / (exp(v1) + exp(v2) + exp(v4) + exp(v6))
df$prob_4 <- exp(v4) / (exp(v1) + exp(v2) + exp(v4) + exp(v6))
df$prob_6 <- exp(v6) / (exp(v1) + exp(v2) + exp(v4) + exp(v6))
```

```{r}
c(mean(df$prob_1),mean(df$prob_2),mean(df$prob_4),mean(df$prob_6))
```
```{r}
v7 <- 1.3413546 + alpha*log(df$TC4*1.20) + beta*(df$TT4*1.20)
```

```{r}
df$prob_1 <- exp(v1) / (exp(v1) + exp(v2) + exp(v4) + exp(v6) + exp(v7))
df$prob_2 <- exp(v2) / (exp(v1) + exp(v2) + exp(v4) + exp(v6) + exp(v7))
df$prob_4 <- exp(v4) / (exp(v1) + exp(v2) + exp(v4) + exp(v6) + exp(v7))
df$prob_6 <- exp(v6) / (exp(v1) + exp(v2) + exp(v4) + exp(v6) + exp(v7))
df$prob_7 <- exp(v7) / (exp(v1) + exp(v2) + exp(v4) + exp(v6) + exp(v7))
c(mean(df$prob_1),mean(df$prob_2),mean(df$prob_4),mean(df$prob_6),mean(df$prob_7))
```


```{r}
LL_1 <- -4240.9
LL_3 <- -4241.5
LL_4 <- -4237.4
AIC_1 <- 2 * 7 - 2 * LL_1
AIC_3 <- 2 * 7 - 2 * LL_3
AIC_4 <- 2 * 8 - 2 * LL_4
c(AIC_1,AIC_3,AIC_4)
```
We find the lowest AIC value for model 4. This is a nesting structure with *car* and *walking* in one nest and *bike* and *public transport* in the other. 


First we find the utility functions
```{r}
df <- read_xlsx('data/PFIII_ex1_2020_data.xlsx')
df$Gender <- df$Gender - 1
coef_time <- -0.0205695
coef_logcost <- -0.3153978
coef_fem <- 0.7133131
coef_nest_bikepub <- 0.6319911
coef_nest_carwalk <- 0.7340071
v1 <- coef_time*df$TT1 + coef_logcost*log(df$TC1)
v2 <- -0.1822955 + coef_time*df$TT2 + coef_logcost*log(df$TC2)
v4 <- 0.9418312 + coef_time*df$TT4 + coef_logcost*log(df$TC4)
v6 <- -1.1596390 + coef_time*df$TT6 + coef_logcost*log(df$TC6) + coef_fem*df$Gender 
```

Then we calculate the within-nest probabilities. This time we have to divide the utility functions with the nest coefficients.
```{r}
logsum_carwalk <- (exp(v1/coef_nest_carwalk) + exp(v4/coef_nest_carwalk))
logsum_bikepub <- (exp(v2/coef_nest_bikepub) + exp(v6/coef_nest_bikepub))
df$prob_1_conditional <- exp(v1/coef_nest_carwalk) / logsum_carwalk
df$prob_4_conditional <- exp(v4/coef_nest_carwalk) / logsum_carwalk
c(mean(df$prob_1_conditional),mean(df$prob_4_conditional))

df$prob_2_conditional <- exp(v2/coef_nest_bikepub) / logsum_bikepub
df$prob_6_conditional <- exp(v6/coef_nest_bikepub) / logsum_bikepub
c(mean(df$prob_2_conditional), mean(df$prob_6_conditional))
```

Then we calculate the probability of choosing a nest
```{r}
df$nestprob_carwalk <- exp(coef_nest_carwalk * log(logsum_carwalk)) / (exp(coef_nest_carwalk * log(logsum_carwalk)) + exp(coef_nest_bikepub * log(logsum_bikepub)))

df$nestprob_bikepub <- exp(coef_nest_bikepub * log(logsum_bikepub)) / (exp(coef_nest_carwalk * log(logsum_carwalk)) + exp(coef_nest_bikepub * log(logsum_bikepub)))
c(mean(df$nestprob_carwalk), mean(df$nestprob_bikepub))
```

And the overall probabilities of choosing one option becomes the within nest probability multiplied by the probability of choosing the nest.
```{r}
df$prob_1 <- df$prob_1_conditional * df$nestprob_carwalk
df$prob_4 <- df$prob_4_conditional * df$nestprob_carwalk

df$prob_2 <- df$prob_2_conditional * df$nestprob_bikepub
df$prob_6 <- df$prob_6_conditional * df$nestprob_bikepub
c(mean(df$prob_1) , mean(df$prob_2) , mean(df$prob_4) , mean(df$prob_6))
```

```{r}
df$VOT_walk <- coef_time / (coef_logcost / log(df$TC1))
df$VOT_bike <- coef_time / (coef_logcost / log(df$TC2))
df$VOT_car <- coef_time / (coef_logcost / log(df$TC4))
df$VOT_pub <- coef_time / (coef_logcost / log(df$TC6))
c(mean(df$VOT_walk),mean(df$VOT_bike),mean(df$VOT_car),mean(df$VOT_pub))
```
```{r}
sum(df$VOT_walk * df$prob_1) / sum(df$prob_1)
sum(df$VOT_bike * df$prob_2) / sum(df$prob_2)
sum(df$VOT_car * df$prob_4) / sum(df$prob_4)
sum(df$VOT_pub * df$prob_6) / sum(df$prob_6)
```

```{r}
#direct
df$elasticity_walk_cost_direct <- coef_logcost*(1-df$prob_1)
df$elasticity_bike_cost_direct <- coef_logcost*(1-df$prob_2)
df$elasticity_car_cost_direct <- coef_logcost*(1-df$prob_4)
df$elasticity_pub_cost_direct <- coef_logcost*(1-df$prob_6)

df$elasticity_walk_time_direct <- df$TT1*coef_time*(1-df$prob_1)
df$elasticity_bike_time_direct <- df$TT2*coef_time*(1-df$prob_2)
df$elasticity_car_time_direct <- df$TT4*coef_time*(1-df$prob_4)
df$elasticity_pub_time_direct <- df$TT6*coef_time*(1-df$prob_6)

#cross
df$elasticity_walk_cost_cross <- -coef_logcost*df$prob_1
df$elasticity_bike_cost_cross <- -coef_logcost*df$prob_2
df$elasticity_car_cost_cross <- -coef_logcost*df$prob_4
df$elasticity_pub_cost_cross <- -coef_logcost*df$prob_6

df$elasticity_walk_time_cross <- -coef_time*df$TT1*df$prob_1
df$elasticity_bike_time_cross <- -coef_time*df$TT2*df$prob_2
df$elasticity_car_time_cross <- -coef_time*df$TT4*df$prob_4
df$elasticity_pub_time_cross <- -coef_time*df$TT6*df$prob_6
```


Weighted averages
```{r}
#direct
sum(df$elasticity_walk_cost_direct * df$prob_1) / sum(df$prob_1)
sum(df$elasticity_bike_cost_direct * df$prob_2) / sum(df$prob_2)
sum(df$elasticity_car_cost_direct * df$prob_4) / sum(df$prob_4)
sum(df$elasticity_pub_cost_direct * df$prob_6) / sum(df$prob_6)

sum(df$elasticity_walk_time_direct * df$prob_1) / sum(df$prob_1)
sum(df$elasticity_bike_time_direct * df$prob_2) / sum(df$prob_2)
sum(df$elasticity_car_time_direct * df$prob_4) / sum(df$prob_4)
sum(df$elasticity_pub_time_direct * df$prob_6) / sum(df$prob_6)

#cross
sum(df$elasticity_walk_cost_cross * df$prob_1) / sum(df$prob_1)
sum(df$elasticity_bike_cost_cross * df$prob_2) / sum(df$prob_2)
sum(df$elasticity_car_cost_cross * df$prob_4) / sum(df$prob_4)
sum(df$elasticity_pub_cost_cross * df$prob_6) / sum(df$prob_6)

sum(df$elasticity_walk_time_cross * df$prob_1) / sum(df$prob_1)
sum(df$elasticity_bike_time_cross * df$prob_2) / sum(df$prob_2)
sum(df$elasticity_car_time_cross * df$prob_4) / sum(df$prob_4)
sum(df$elasticity_pub_time_cross * df$prob_6) / sum(df$prob_6)
```



```{r}
market_share_walk_base <- mean(df$prob_1)*nrow(df)
market_share_bike_base <- mean(df$prob_2)*nrow(df)
market_share_car_base <- mean(df$prob_4)*nrow(df)
market_share_pub_base <- mean(df$prob_6)*nrow(df)

df2 <- df

v1 <- coef_time*df2$TT1 + coef_logcost*log(df2$TC1)
v2 <- -0.1822955 + coef_time*df2$TT2 + coef_logcost*log(df2$TC2)
v4 <- 0.9418312 + coef_time*df2$TT4 + coef_logcost*log(df2$TC4*1.10)
v6 <- -1.1596390 + coef_time*df2$TT6 + coef_logcost*log(df2$TC6) + coef_fem*df2$Gender 

logsum_carwalk <- (exp(v1/coef_nest_carwalk) + exp(v4/coef_nest_carwalk))
logsum_bikepub <- (exp(v2/coef_nest_bikepub) + exp(v6/coef_nest_bikepub))
df2$prob_1_conditional <- exp(v1/coef_nest_carwalk) / logsum_carwalk
df2$prob_4_conditional <- exp(v4/coef_nest_carwalk) / logsum_carwalk

df2$prob_2_conditional <- exp(v2/coef_nest_bikepub) / logsum_bikepub
df2$prob_6_conditional <- exp(v6/coef_nest_bikepub) / logsum_bikepub

df2$nestprob_carwalk <- exp(coef_nest_carwalk * log(logsum_carwalk)) / (exp(coef_nest_carwalk * log(logsum_carwalk)) + exp(coef_nest_bikepub * log(logsum_bikepub)))

df2$nestprob_bikepub <- exp(coef_nest_bikepub * log(logsum_bikepub)) / (exp(coef_nest_carwalk * log(logsum_carwalk)) + exp(coef_nest_bikepub * log(logsum_bikepub)))

df2$prob_1 <- df2$prob_1_conditional * df2$nestprob_carwalk
df2$prob_4 <- df2$prob_4_conditional * df2$nestprob_carwalk

df2$prob_2 <- df2$prob_2_conditional * df2$nestprob_bikepub
df2$prob_6 <- df2$prob_6_conditional * df2$nestprob_bikepub

market_share_walk_scenario <- mean(df2$prob_1)*nrow(df2)
market_share_bike_scenario <- mean(df2$prob_2)*nrow(df2)
market_share_car_scenario <- mean(df2$prob_4)*nrow(df2)
market_share_pub_scenario <- mean(df2$prob_6)*nrow(df2)

((market_share_walk_scenario - market_share_walk_base)/ market_share_walk_base) / 0.10
((market_share_bike_scenario - market_share_bike_base)/ market_share_bike_base) / 0.10
((market_share_car_scenario  - market_share_car_base) / market_share_car_base)/ 0.10
((market_share_pub_scenario  - market_share_pub_base) / market_share_pub_base)/ 0.10
```



```{r}
v1 <- coef_time*df2$TT1 + coef_logcost*log(df2$TC1)
v2 <- -0.1822955 + coef_time*df2$TT2 + coef_logcost*log(df2$TC2)
v4 <- 0.9418312 + coef_time*df2$TT4*1.10 + coef_logcost*log(df2$TC4)
v6 <- -1.1596390 + coef_time*df2$TT6 + coef_logcost*log(df2$TC6) + coef_fem*df2$Gender 

logsum_carwalk <- (exp(v1/coef_nest_carwalk) + exp(v4/coef_nest_carwalk))
logsum_bikepub <- (exp(v2/coef_nest_bikepub) + exp(v6/coef_nest_bikepub))
df2$prob_1_conditional <- exp(v1/coef_nest_carwalk) / logsum_carwalk
df2$prob_4_conditional <- exp(v4/coef_nest_carwalk) / logsum_carwalk

df2$prob_2_conditional <- exp(v2/coef_nest_bikepub) / logsum_bikepub
df2$prob_6_conditional <- exp(v6/coef_nest_bikepub) / logsum_bikepub

df2$nestprob_carwalk <- exp(coef_nest_carwalk * log(logsum_carwalk)) / (exp(coef_nest_carwalk * log(logsum_carwalk)) + exp(coef_nest_bikepub * log(logsum_bikepub)))

df2$nestprob_bikepub <- exp(coef_nest_bikepub * log(logsum_bikepub)) / (exp(coef_nest_carwalk * log(logsum_carwalk)) + exp(coef_nest_bikepub * log(logsum_bikepub)))

df2$prob_1 <- df2$prob_1_conditional * df2$nestprob_carwalk
df2$prob_4 <- df2$prob_4_conditional * df2$nestprob_carwalk

df2$prob_2 <- df2$prob_2_conditional * df2$nestprob_bikepub
df2$prob_6 <- df2$prob_6_conditional * df2$nestprob_bikepub

market_share_walk_scenario <- mean(df2$prob_1)*nrow(df2)
market_share_bike_scenario <- mean(df2$prob_2)*nrow(df2)
market_share_car_scenario <- mean(df2$prob_4)*nrow(df2)
market_share_pub_scenario <- mean(df2$prob_6)*nrow(df2)

((market_share_walk_scenario - market_share_walk_base)/ market_share_walk_base) / 0.10
((market_share_bike_scenario - market_share_bike_base)/ market_share_bike_base) / 0.10
((market_share_car_scenario  - market_share_car_base) / market_share_car_base)/ 0.10
((market_share_pub_scenario  - market_share_pub_base) / market_share_pub_base)/ 0.10
```


```{r}


calculate_ms <- function(asc){
  v1 <- asc[1] + coef_time*df2$TT1 + coef_logcost*log(df2$TC1)
  v2 <- asc[2] + coef_time*df2$TT2 + coef_logcost*log(df2$TC2)
  v4 <- asc[3] + coef_time*df2$TT4 + coef_logcost*log(df2$TC4)
  v6 <- asc[4] + coef_time*df2$TT6 + coef_logcost*log(df2$TC6) + coef_fem*df2$Gender 
  v7 <- asc[5] + coef_time*df2$TT4*1.20 + coef_logcost*log(df2$TC4*1.20)
    
  logsum_carwalk <- (exp(v1/coef_nest_carwalk) + exp(v4/coef_nest_carwalk) + exp(v7/coef_nest_carwalk))
  logsum_bikepub <- (exp(v2/coef_nest_bikepub) + exp(v6/coef_nest_bikepub))
  df2$prob_1_conditional <- exp(v1/coef_nest_carwalk) / logsum_carwalk
  df2$prob_4_conditional <- exp(v4/coef_nest_carwalk) / logsum_carwalk
  df2$prob_7_conditional <- exp(v7/coef_nest_carwalk) / logsum_carwalk
  
  df2$prob_2_conditional <- exp(v2/coef_nest_bikepub) / logsum_bikepub
  df2$prob_6_conditional <- exp(v6/coef_nest_bikepub) / logsum_bikepub
  
  df2$nestprob_carwalk <- exp(coef_nest_carwalk * log(logsum_carwalk)) / (exp(coef_nest_carwalk * log(logsum_carwalk)) + exp(coef_nest_bikepub * log(logsum_bikepub)))
  
  df2$nestprob_bikepub <- exp(coef_nest_bikepub * log(logsum_bikepub)) / (exp(coef_nest_carwalk * log(logsum_carwalk)) + exp(coef_nest_bikepub * log(logsum_bikepub)))
  
  df2$prob_1 <- df2$prob_1_conditional * df2$nestprob_carwalk
  df2$prob_4 <- df2$prob_4_conditional * df2$nestprob_carwalk
  df2$prob_7 <- df2$prob_7_conditional * df2$nestprob_carwalk
  
  df2$prob_2 <- df2$prob_2_conditional * df2$nestprob_bikepub
  df2$prob_6 <- df2$prob_6_conditional * df2$nestprob_bikepub
  
  return(c(mean(df2$prob_1),mean(df2$prob_2),mean(df2$prob_4),mean(df2$prob_6),mean(df2$prob_7)))
}

calibrate_ms <- function(MS_est,MS_given,asc){
  asc[1] <- asc[1] + log(MS_given[1]/MS_est[1])
  asc[2] <- asc[2] + log(MS_given[2]/MS_est[2])
  asc[3] <- asc[3] + log(MS_given[3]/MS_est[3])
  asc[4] <- asc[4] + log(MS_given[4]/MS_est[4])
  asc[5] <- asc[5] + log(MS_given[5]/MS_est[5])
  return(asc)
}
  
```

k=0
```{r}
MS_given <- c(0.15,0.25,0.39,0.08,0.13)
asc <- c(0,-0.1822955,0.9418312,-1.1596390,0.9418312)
MS_est <- calculate_ms(asc)
MS_est

```
k=1
```{r}
asc <- calibrate_ms(MS_est,MS_given,asc)
asc
```
```{r}
MS_est <- calculate_ms(asc)
MS_est
```
k=2
```{r}
asc <- calibrate_ms(MS_est,MS_given,asc)
asc
```

```{r}
MS_est <- calculate_ms(asc)
MS_est
```
k=3
```{r}
asc <- calibrate_ms(MS_est,MS_given,asc)
asc
```
```{r}
MS_est <- calculate_ms(asc)
MS_est
```
k=4
```{r}
asc <- calibrate_ms(MS_est,MS_given,asc)
asc
```

```{r}
MS_est <- calculate_ms(asc)
MS_est
```
k=5
```{r}
asc <- calibrate_ms(MS_est,MS_given,asc)
asc
```

```{r}
MS_est <- calculate_ms(asc)
MS_est
```
k=6
```{r}
asc <- calibrate_ms(MS_est,MS_given,asc)
asc
```

```{r}
MS_est <- calculate_ms(asc)
MS_est
```
k=6
```{r}
asc <- calibrate_ms(MS_est,MS_given,asc)
asc
```

```{r}
MS_est <- calculate_ms(asc)
MS_est
```
k=7
```{r}
asc <- calibrate_ms(MS_est,MS_given,asc)
asc
```

```{r}
MS_est <- calculate_ms(asc)
MS_est
```
```{r}
calculate_ms_sc <- function(asc){
  v1 <- asc[1] + coef_time*df2$TT1 + coef_logcost*log(df2$TC1)
  v2 <- asc[2] + coef_time*df2$TT2 + coef_logcost*log(df2$TC2)
  v4 <- asc[3] + coef_time*df2$TT4 + coef_logcost*log(df2$TC4)
  v6 <- asc[4] + coef_time*df2$TT6 + coef_logcost*log(df2$TC6) + coef_fem*df2$Gender 
  v7 <- asc[5] + coef_time*df2$TT4 + coef_logcost*log(df2$TC4)
    
  logsum_carwalk <- (exp(v1/coef_nest_carwalk) + exp(v4/coef_nest_carwalk) + exp(v7/coef_nest_carwalk))
  logsum_bikepub <- (exp(v2/coef_nest_bikepub) + exp(v6/coef_nest_bikepub))
  df2$prob_1_conditional <- exp(v1/coef_nest_carwalk) / logsum_carwalk
  df2$prob_4_conditional <- exp(v4/coef_nest_carwalk) / logsum_carwalk
  df2$prob_7_conditional <- exp(v7/coef_nest_carwalk) / logsum_carwalk
  
  df2$prob_2_conditional <- exp(v2/coef_nest_bikepub) / logsum_bikepub
  df2$prob_6_conditional <- exp(v6/coef_nest_bikepub) / logsum_bikepub
  
  df2$nestprob_carwalk <- exp(coef_nest_carwalk * log(logsum_carwalk)) / (exp(coef_nest_carwalk * log(logsum_carwalk)) + exp(coef_nest_bikepub * log(logsum_bikepub)))
  
  df2$nestprob_bikepub <- exp(coef_nest_bikepub * log(logsum_bikepub)) / (exp(coef_nest_carwalk * log(logsum_carwalk)) + exp(coef_nest_bikepub * log(logsum_bikepub)))
  
  df2$prob_1 <- df2$prob_1_conditional * df2$nestprob_carwalk
  df2$prob_4 <- df2$prob_4_conditional * df2$nestprob_carwalk
  df2$prob_7 <- df2$prob_7_conditional * df2$nestprob_carwalk
  
  df2$prob_2 <- df2$prob_2_conditional * df2$nestprob_bikepub
  df2$prob_6 <- df2$prob_6_conditional * df2$nestprob_bikepub
  
  return(c(mean(df2$prob_1),mean(df2$prob_2),mean(df2$prob_4),mean(df2$prob_6),mean(df2$prob_7)))
}
```

```{r}
MS_est <- calculate_ms_sc(asc)
MS_est
```


